---
title: "Atalia_Horenshtien_Covid"
author: "Atalia Horenshtien"
date: "3/6/2021"
output:
  word_document: default
  html_document: default
---

# setup and load libraries

```{r setup, include=FALSE}
##load libraries
library(tidyverse)
```

```{r, include=FALSE}
### working directory
setwd("~/Documents/MsBA/Data strategy/class5")
```

```{r}
##load the data
covid_data = read.csv('owid-covid-data.csv')
```

```{r}
colnames(covid_data)
```

# Layer 0:

## Question: Who are the top 5 coutries that are doing the best in vaccination?

## Graphic:
```{r}
#filter data for the recent record ('2021-02-28'), calculate the percentage_vaccinated and pick the top 5 coutries
covid_data %>% 
  filter(date == '2021-02-28') %>% 
  mutate(percentage_vaccinated = people_vaccinated/population *100) %>% 
  filter(percentage_vaccinated != 'NA') %>% 
  select(location, percentage_vaccinated) %>% 
  arrange(desc(percentage_vaccinated)) %>% 
  head(n=5)
```
## Findings: Israel is doing the best with 54.57% of the population is being vaccinated as for Feb 28, 2021.

I'm from Israel, therefore I'll continue my research on Israel data.
```{r}
##slice data for Israel
covid_data_Israel = covid_data %>% 
  filter(location == 'Israel')
```

# Layer 1:

## Question: Is total deaths increase when total cases increase?

## Graphic:

```{r}

covid_data_Israel %>%
  ggplot(data = ., aes(x = total_cases, y = total_deaths)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title="sum of total cases per million vs stringency index")

```
## Findings: We can see that as total cases increase the total deaths increase as well. lets see if the time in lockdown has any affect on this. 

# Layer 2:

## Question: Is Israel Lockdown has a success? is being more time in lockdown reduce the total cases and/or total deaths per million?

## Graphic:

```{r}

covid_data_Israel %>%
  ggplot(data = ., aes(x = total_cases_per_million, y = stringency_index)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title="sum of total cases per million vs stringency index")

```


```{r}

covid_data_Israel %>%
  ggplot(data = ., aes(x = total_deaths_per_million, y = stringency_index)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title="sum of total deaths per million vs stringency index")

```
## Findings: Lockdown in Israel is not brining a significant success, we can see that there is a positive relationship between the stringency index and total cases per million and between the stringency index and total deathes per million. now that more than 50% of the popultion in the country is vaccinated, let's see if has any effect on the new cases 


# Layer 3:

## Question: Is vaccinating people helps with stop spreading the virus and reduce the new cases?

## Graphic:

```{r}

covid_data_Israel %>%
  ggplot(data = ., aes(x = people_vaccinated, y = new_cases)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title="new cases as people getting vaccinated")

```
## Findings: We can see that the as people getting more vaccinated the number of new cases decline. it make a lot of sense because vaccinated people are less to get the virus. 
Israel started vaccination in middle of Dec 2020, therefore many people already got the second shot and are fully vaccinated. 71.31% of the vaccinated people are fully vaccinated. Let's see if it has an effect on the positive rate.

```{r}
covid_data_Israel %>% 
  filter(date == '2021-02-28') %>% 
  summarize(people_vaccinated, people_fully_vaccinated, percentage_fully_vaccinated = people_fully_vaccinated/people_vaccinated)

```

# Layer 4:

## Question: Is a population with fully vaccinated people has a high or low positive rate?

## Graphic:
```{r}

covid_data_Israel %>%
  ggplot(data = ., aes(x = people_fully_vaccinated, y = positive_rate)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title="positive rate as people getting fully vaccinated")

```
## Findings: We can see that the positive rate is declining as more people are getting fully vaccinated.
